Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-12 chapters/lit_review_new.tex
Changes 1. Reformatting for intro, lit review and methods complete

The thermodynamic equilibrium of a binary fluid mixture can be described by its Gibbs free energy, given as $G = \Delta H - T \Delta S$, where \(\Delta H\) is the enthalpy change, 
\(T\) is temperature, and \(\Delta S\) is the entropy change. The onset of phase separation occurs at the binodal line, defined where \(\frac{dG}{dx} = 0\), with \(x\) representing the composition. 
Although the system is metastable, it remains uniform until fluctuations give rise to small nuclei. The free energy of each nucleus reflects a 
balance between the bulk free energy gain from phase separation and the interfacial energy penalty, expressed as \(\Delta G = \frac{4}{3}\pi r^3 \Delta g_v + 4 \pi r^2 \sigma\), where \(r\) is 
the nucleus radius, \(\Delta g_v\) is the free energy change per unit volume, and \(\sigma\) is the interfacial tension \cite{thanh_mechanisms_2014}. 
Nuclei that exceed the critical radius \(r_c = \frac{2\sigma}{|\Delta g_v|}\) will grow, driving the system toward phase separation via nucleation and growth \cite{thanh_mechanisms_2014}. 
This transformation process is commonly modeled by the Avrami equation, \(Y = 1 - \exp(-K t^n)\), where \(Y\) is the 
transformed fraction, \(K\) is a rate constant incorporating nucleation and growth kinetics, and \(n\) is the Avrami exponent, which reflects the dimensionality and mechanism of the transformation
\cite{avrami_kinetics_1939} .

At the binodal, the system is still metastable, meaning that unless perturbations are introduced to form nuclei, the mixture remains homogeneous \cite{thanh_mechanisms_2014}.
However, a deeper quench into the two-phase region leads to the spinodal line, defined by the stability criterion \(\frac{d^2G}{dx^2} = 0\) \cite{cahn_free_1958}. 
Within this region, the system is unstable to even infinitesimal composition fluctuations, and phase 
separation proceeds spontaneously without the need for nucleation. This process, known as spinodal decomposition, is characterized by the continuous and simultaneous growth of interconnected 
domains of both phases \cite{cahn_free_1958}. Unlike nucleation and growth, which relies on localized events, spinodal decomposition 
results in a bicontinuous structure driven by the system's attempt to minimize 
free energy through composition modulation \cite{cahn_free_1958}. The dynamics of spinodal decomposition were characterized by Cahn and Hilliard \cite{cahn_free_1958} . 
We first begin by defining the free energy of a regular solution, known as the bulk free energy.

\begin{equation}
    \Delta G_m = \chi (1 - \phi)\phi + N_a k_B T \left[(1 - \phi)\ln(1 - \phi) + \phi \ln(\phi)\right]
\end{equation}

@@

\begin{equation}
    F[\phi] = \int \left( f(\phi) + \frac{\kappa}{2}|\nabla \phi|^2 \right) \, dV
\end{equation}

@@

\begin{equation}
    \mu = \frac{\delta F}{\delta \phi} = \frac{df}{d\phi} - \kappa \nabla^2 \phi
\end{equation}

@@

where $M$ is the mobility and could potentially be a function of $\phi$. Assuming no advection and applying mass conservation for the order parameter, we obtain,

\begin{equation}
    \frac{\partial \phi}{\partial t} = -\nabla \cdot \vec{J}
\end{equation}

Substituting in the expression for \(\vec{J}\) leads to the Cahn-Hilliard equation \cite{cahn_free_1958},

\begin{equation}
    \frac{\partial \phi}{\partial t} = \nabla \cdot \left( M \nabla \mu \right)
\end{equation}

This equation describes the diffusive evolution of the conserved field $\phi$ during phase separation. It captures the system's tendency to minimize its total free energy 
by redistributing the order parameter according to gradients in chemical potential. The Cahn-Hilliard framework naturally incorporates interfacial tension and allows for smooth, 
continuous transitions between phases, making it highly suitable for modeling non-ideal mixing in systems such as bijels.

\textcolor{blue}{add surface tension calculation}

\textcolor{blue}{add interface width calculation}

@@

When micron sized particles are adsorbed by interfaces the attachment energy is large enough that they can become irreversibly trapped at fluid-fluid interfaces \cite{ngai_particle-stabilized_2015}. 
Once adsorbed, these 
particles reduce interfacial tension and act as physical barriers that prevent the coalescence of droplets, thereby making them highly effective emulsion stabilizers. Unlike molecular surfactants, 
which are dynamic and can desorb under certain conditions, colloidal particles remain adsorbed at the interface due to the significant energy required to detach them \cite{ngai_particle-stabilized_2015}. 
This mechanism 
is especially useful in systems such as Pickering emulsions, where the interfacial coverage and wettability of the particles determine the type and stability of the emulsion formed 
\cite{ngai_particle-stabilized_2015,velankar_non-equilibrium_2015}.
As such, understanding particle behavior at interfaces is essential for controlling the structure and longevity of particle-stabilized systems \cite{ngai_particle-stabilized_2015}. 
For particles that are smooth and simple and at an interface of oil and water, the attachment energy at the fluid interface can be calculated the Pieranski model, 
$E_{ads} = \sigma A_{rm} (1 - \cos{\theta_c})^2$, where $E_{ads}$ is the free energy reduction upon particle adsorption, $\sigma$ is the surface tension between the two 
fluids, and $\theta_c$ is the contact angle of the stabilizing particle.

This expression is derived from an energy balance between the interactions between the interface and the 
fluid phases, with a schematic shown in Figure \ref{fig:pieranski_model},

@@

From the Pieranski model, the adsorption energy can be seen to depend on the contact angle, surface tension and contact area of the particles. This indicates that larger, neutrally wetting
particles adsorbing onto interfaces with high surface tension have a high driving force for adsorption. \cite{ngai_particle-stabilized_2015,reeves_particle-size_2015} This adsorption process 
can be orders of magnitude larger than the thermal energy of the system, calculated as $k_b T$ where $T$ is the temperature of the system and $k_b$ is the Boltzmann constant. This results in 
irreversible adsorption of particles at interfaces \cite{binks_pickering_2001}.

In addition to the adsorption of particles on the interface, the contact angle of the particles determine the dispersed phase of an emulsion and its microstructure through the
Bancroft rule \cite{velankar_non-equilibrium_2015,ngai_particle-stabilized_2015}. 
Hydrophillic particles tend to prefer to form oil in water emulsions while hydrophobic particles prefer to form water in oil emulsions \cite{ngai_particle-stabilized_2015}.
This occurs as the particles preferentially wet one phase over another, causing the interface to bend as the emulsion forms. Neutrally wetting particles are preferred in the formation of 
bijels as no preferential curvature is imparted onto the bijels, fulfilling one of the requirements of a minimal surface in that the mean curvature must be zero \cite{jinnai_interfacial_2001}.

% from a force 
% balance between the force imparted from the particle on each fluid and the fluids between one another, known as the Young-Dupre equation.

% \begin{equation}
%     \cos{\theta_c} = \frac{\sigma_{po} - \sigma_{pw}}{\sigma_{ow}}
% \end{equation}

% Where $\theta_c$ is the contact angle, $\sigma_{po}$ is the surface tension between particle and oil, $\sigma_{pw}$ is the surface tension between particle and water and
% $\sigma_{ow}$ is the surface tension between oil and water. The particle is considered to be preferentially wet by water or hydrophilic if $\theta_c < 90 ^{\circ}$ or 
% hydrophobic if $\theta_c > 90 ^{\circ}$. The particle is said to be neutrally wetting if $\theta_c = 90 ^{\circ}$.

% \begin{figure}
%     \centering
%     \includegraphics[scale = 0.5]{../figures/literature_review/particle_at_interface.png }
%     \caption{Particle schematic at an interface demonstrating the various quantities of interest. Reprinted Figure 2.1 from
%              Particle-Stabilized Emulsions and Colloids: Formation and Applications; Ngai, T., Bon, S. A. F., Eds.; RSC soft matter series; 
%              with permission Royal Society of Chemistry, RSC Publishing Cambridge, 2015 under license number 1597404-1}
%     \label{fig:pieranski_model}
% \end{figure}

% The energy of the system before and after particle adsorption is different and can be characterized for a simple system with parameters shown in Figure \ref{fig:pieranski_model},

% \begin{equation}
%     \Delta E_{IW} = E_I - E_w
% \end{equation}

% Where $\Delta E_{I}$ is the energy of the system when the particle is in the interface and $\Delta E_W$ is the energy of the system when the particle is completely in the water phase.
% We define $\Delta E_{I}$ as

% \begin{equation}
%     E_I = A_w \sigma_{pw} + A_o\sigma_{po} + A_{I_1}\sigma_{ow} + V_w P_w + V_o P_o
% \end{equation}

% Where $A_w$ and $A_o$ are the surface areas of the particle in the water and oil phases respectively, $V_w$ and $V_o$ are the volumes of water and oil phases respectively and 
% $P_w$ and $P_o$ are the pressures of water and oil phases respectively. $A_{I_1}$ is the area of the interface with the particle. 
% A similar equation can be written for the energy of the water phase, $E_w$

% \begin{equation}
%     E_w = (A_w + A_o)\sigma_{pw} + A_{I_2}\sigma_{ow} + (V_w + V_o)P_w
% \end{equation}

% $A_{I_2}$ is the area of the interface without the particle. The change in the interface energy upon adsorption of the particle can be computed to obtain

% \begin{equation}
%     \Delta E_{IW} = A_o(\sigma_{po} - \sigma_{pw}) - A_p\sigma_{ow} + V_o(P_o - P_w)
% \end{equation}

% Substituting the Young-Laplace law for a spherical droplet, $P_w - P_o = \frac{2\sigma_{ow}}{R_d}$ where $R_d$ is the droplet size, for $(P_o - P_w)$ and the Young-Dupre contact angle
% expression, the following is obtained

% \begin{equation}
%     \Delta E_{IW} = \sigma_{ow}(A_o\cos{\theta_c} - A_p - \frac{2 V_o}{R_d})
% \end{equation}

% A similar construction can be done for the oil phase to obtain

% \begin{equation}
%     \Delta E_{IW} = -\sigma_{ow}(A_w\cos{\theta_c} + A_p - \frac{2 V_o}{R_d})
% \end{equation}

% After making appropriate geometrical substitutions for $V_o$ and $A_p$, the Piersanski model is recovered,

% \begin{equation}
%     \Delta E_{ads} = \pi r^2 \sigma_{ow} (1 - |\cos{\theta_c}|)^2
% \end{equation}

% From this expression, the adsorption energy can be seen to depend on the contact angle, surface tension and contact area of the particles. This indicates that larger, neutrally wetting
% particles adsorbing onto interfaces with high surface tension have a high driving force for adsorption. \cite{ngai_particle-stabilized_2015,reeves_particle-size_2015} This adsorption process 
% can be orders of magnitude larger than the thermal energy of the system, calculated as $k_b T$ where $T$ is the temperature of the system and $k_b$ is the Boltzmann constant. This results in 
% irreversible adsorption of particles at interfaces. \cite{binks_pickering_2001}

% In addition to the adsorption of particles on the interface, the contact angle of the particles determine the dispersed phase of an emulsion and its microstructure. 
% \cite{velankar_non-equilibrium_2015}
% Hydrophillic particles tend to prefer to form oil in water emulsions while hydrophobic particles prefer to form water in oil emulsions. \cite{ngai_particle-stabilized_2015}
% This occurs as the particles preferentially wet one phase over another, causing the interface to bend as the emulsion forms. Neutrally wetting particles are preferred in the formation of 
% bijels as no preferential curvature is imparted onto the bijels, fulfilling one of the requirements of a minimal surface in that the mean curvature must be zero. \cite{jinnai_interfacial_2001}

@@

In contrast, explicit particle models represent colloids as finite-size rigid bodies with well-defined shape, volume, and surface boundary conditions. 
\cite{ladd_numerical_1994, jansen_bijels_2011,gunther_timescales_2014}
These particles interact 
with the surrounding fluid through no-slip or slip boundary conditions, and their motion responds dynamically to hydrodynamic, capillary, and interparticle forces. By resolving 
particle geometry, these models allow for accurate representation of rotational motion, surface wetting and interfacial adsorption.
\cite{jansen_bijels_2011, gunther_lattice_2013}
This is important in bijel systems, where particle-induced jamming depend on particle shape, size, and surface wetting.

An intermediate approach between point particles and explicitly resolved rigid bodies is the immersed boundary method (IBM), which represents extended particles as collections of 
point markers embedded in the fluid. \cite{peskin_immersed_2002, luo_immersed-boundary_2008, spandan_parallel_2017}
These markers interact with the fluid through momentum exchange that impose motion onto the surrounding fluid, effectively reconstructing 
the dynamics of a deformable or rigid object without requiring an explicit boundary mesh. \cite{peskin_immersed_2002, luo_immersed-boundary_2008, spandan_parallel_2017} 
IBM has been widely used in biological and colloidal systems, to model the dynamics of deformable bodies. \cite{peskin_immersed_2002, luo_immersed-boundary_2008, spandan_parallel_2017}

For systems in which interfacial structure, jamming, or near-field hydrodynamics play a central role as in bijel formation, explicit particle modeling 
is generally preferred as we are able to resolve the interfacial structure and near field hydrodynamics without the complications of IBM.

@@

Magnetically responsive particles can generally be classified as paramagnetic or ferromagnetic, depending on the nature and strength of their magnetization in response to external 
magnetic fields. Paramagnetic particles do not possess a permanent magnetic moment but acquire a reversible magnetization that aligns with the applied field and 
vanishes once the field is removed. \cite{sinn_magnetically_2011} Their response is typically linear with field strength and does not involve magnetic hysteresis. 
In contrast, ferromagnetic particles exhibit strong, permanent magnetization due to the alignment of magnetic domains even in the absence of an external field. These particles can 
exhibit hysteresis and significantly stronger dipole-dipole interactions. \cite{sinn_magnetically_2011}
As a result, ferromagnetic particles are more likely to form irreversible aggregates or chains, while paramagnetic particles 
respond more gently and reversibly, making them preferable in systems where tunability and field-off stability are important. The choice between these types depends on the desired 
degree of responsiveness, reversibility, and control in applications such as field-directed assembly or responsive emulsions.

Magnetic particles are often modeled using the magnetic dipole approximation, which treats each particle as possessing a single point dipole moment \(\vec{m}\) located at its center.
\cite{davies_assembling_2014}
The dipole-dipole interaction between two particles separated by a vector \(\vec{r}\) is described by the potential:

\begin{equation}
    U_{\text{d}} = \frac{\mu_0}{4\pi r^3} \left[ \vec{m}_1 \cdot \vec{m}_2 - 3(\vec{m}_1 \cdot \hat{r})(\vec{m}_2 \cdot \hat{r}) \right]
\end{equation}

This model captures the long-range, anisotropic nature of magnetic interactions and is commonly used due to its simplicity and computational efficiency. It is especially effective when 
particles are small, well-separated, and uniformly magnetized, such that higher-order multipole contributions can be neglected. However, the dipole model becomes less accurate when particles 
aggregate closely or experience strong magnetic fields that cause internal magnetization to deviate from a uniform state.

To address these limitations, the demagnetization tensor method offers a more refined model by considering the particle as a uniformly magnetized ellipsoid and calculating the internal magnetic 
field using an analytical or numerically tabulated demagnetization tensor. \cite{takahashi_ellipsoids_2017} 
This method accounts for the effect of particle shape on the distribution of internal magnetization and enables the modeling of induced magnetic 
dipoles that respond to local fields. The total magnetic field of an ellipsoids can be written,

\begin{equation}
    \mathbf{H}(r) = \mathbf{H}_0 + \mathbf{N}(r)\mathbf{K}\mathbf{H}^{\dagger}
\end{equation}

Where $\mathbf{H}(r)$ is the total magnetic field of the ellipsoids, $\mathbf{H}_0$ is the inducing magnetic field in each direction,
$\mathbf{K}$ is the susceptibility tensor and $\mathbf{H}^{\dagger}$ is the resultant uniform magnetic field at any point within
the ellipsoids. \cite{takahashi_ellipsoids_2017}  This approach is particularly useful for modeling spheroidal or ellipsoidal particles and is widely 
used in magnetorheological systems, chain-forming ferrofluids, and dense suspensions where particle-particle interactions are shape-dependent.

Since bijel formation and stabilization primarily depend on interfacial 
jamming and long-range interactions between particles at fluid-fluid interfaces, the magnetic dipole approximation is often sufficient to model key behaviors such as chaining, orientation under external 
fields, and capillary-mediated assembly. However, if particle shape anisotropy, proximity effects, or induced magnetization in response to local fields are important, incorporating demagnetization tensors 
may offer improved realism.

>>>>

The thermodynamic equilibrium of a binary fluid mixture can be described by its Gibbs free energy, given as $G = \Delta H - T \Delta S$, where \(\Delta H\) is the enthalpy change, 
\(T\) is temperature, and \(\Delta S\) is the entropy change. The onset of phase separation occurs at the binodal line, defined where \(\frac{dG}{dx} = 0\), with \(x\) representing the composition. 
Although the system is metastable, it remains uniform until fluctuations give rise to small nuclei. The free energy of each nucleus reflects a 
balance between the bulk free energy gain from phase separation and the interfacial energy penalty, expressed as \(\Delta G = \frac{4}{3}\pi r^3 \Delta g_v + 4 \pi r^2 \sigma\), where \(r\) is 
the nucleus radius, \(\Delta g_v\) is the free energy change per unit volume, and \(\sigma\) is the interfacial tension \cite{thanh_mechanisms_2014}. 
Nuclei that exceed the critical radius \(r_c = \frac{2\sigma}{|\Delta g_v|}\) will grow, driving the system toward phase separation via nucleation and growth \cite{thanh_mechanisms_2014}. 
This transformation process is commonly modeled by the Avrami equation, \(Y = 1 - \exp(-K t^n)\), where \(Y\) is the 
transformed fraction, \(K\) is a rate constant incorporating nucleation and growth kinetics, and \(n\) is the Avrami exponent, which reflects the dimensionality and mechanism of the transformation
\cite{avrami_kinetics_1939}.

At the binodal, the system is still metastable, meaning that unless perturbations are introduced to form nuclei, the mixture remains homogeneous \cite{thanh_mechanisms_2014}.
However, a deeper quench into the two-phase region leads to the spinodal line, defined by the stability criterion \(\frac{d^2G}{dx^2} = 0\) \cite{cahn_free_1958}. 
Within this region, the system is unstable to even infinitesimal composition fluctuations, and phase 
separation proceeds spontaneously without the need for nucleation. This process, known as spinodal decomposition, is characterized by the continuous and simultaneous growth of interconnected 
domains of both phases \cite{cahn_free_1958}. Unlike nucleation and growth, which relies on localized events, spinodal decomposition 
results in a bicontinuous structure driven by the system's attempt to minimize 
free energy through composition modulation \cite{cahn_free_1958}. The dynamics of spinodal decomposition were characterized by Cahn and Hilliard \cite{cahn_spinodal_1961}. 
We first begin by defining the free energy of a regular solution, known as the bulk free energy given by

\begin{equation}
    G_m (c) = \chi (1 - c)c + N_a k_B T \left[(1 - c)\ln(1 - c) + c \ln(c)\right]
\end{equation}

@@

\begin{equation}
    F[c] = \int \left( f(c) + \frac{\kappa}{2}|\nabla c|^2 \right) \, dV
\end{equation}

@@

\begin{equation}
    \mu = \frac{\delta F}{\delta c} = \frac{df}{dc} - \kappa \nabla^2 c
\end{equation}

@@

where $M$ is the mobility and could potentially be a function of $c$. Assuming no advection and applying mass conservation for the order parameter, we obtain,

\begin{equation}
    \frac{\partial c}{\partial t} = -\nabla \cdot \vec{J}
\end{equation}

Substituting in the expression for \(\vec{J}\) leads to the Cahn-Hilliard equation 18 in reference \cite{cahn_spinodal_1961},

\begin{equation}
    \frac{\partial c}{\partial t} = \nabla \cdot \left( M \nabla \mu \right)
\end{equation}

This equation describes the diffusive evolution of the conserved field $c$ during phase separation. It captures the system's tendency to minimize its total free energy 
by redistributing the order parameter according to gradients in chemical potential. The Cahn-Hilliard framework naturally incorporates interfacial tension and allows for smooth, 
continuous transitions between phases, making it highly suitable for modeling non-ideal mixing in systems such as bijels.

@@

When micron sized particles are adsorbed by interfaces the attachment energy is large enough that they can become irreversibly trapped at fluid-fluid interfaces \cite{ngai_particle-stabilized_2015}. 
Once adsorbed, these 
particles reduce interfacial tension and act as physical barriers that prevent the coalescence of droplets, thereby making them highly effective emulsion stabilizers. Unlike molecular surfactants, 
which are dynamic and can desorb under certain conditions, colloidal particles remain adsorbed at the interface due to the significant energy required to detach them \cite{ngai_particle-stabilized_2015}. 
This mechanism 
is especially useful in systems such as Pickering emulsions, where the interfacial coverage and wettability of the particles determine the type and stability of the emulsion formed 
\cite{ngai_particle-stabilized_2015,velankar_non-equilibrium_2015}.
As such, understanding particle behavior at interfaces is essential for controlling the structure and longevity of particle-stabilized systems \cite{ngai_particle-stabilized_2015}. 
For particles that are smooth and simple and at an interface of oil and water, the attachment energy at the fluid interface can be calculated the Pieranski model, 
$G_{ads} = \sigma A_{rm} (1 - \cos{\theta_c})^2$, where $G_{ads}$ is the free energy reduction upon particle adsorption, $\sigma$ is the surface tension between the two 
fluids, and $\theta_c$ is the contact angle of the stabilizing particle. A schematic of a particle adsorbing at an interface is shown in Figure \ref{fig:pieranski_model},

@@

To obtain the Pieranski model, an energy balance between the interactions between the interface and the fluid phases can be written. The energy of the system before and after particle adsorption is 
different and can be characterized for the particle moving from water into the interface as,

\begin{equation}
    \Delta E_{IW} = E_I - E_w
\end{equation}

Where $\Delta E_{I}$ is the energy of the system when the particle is in the interface and $\Delta E_W$ is the energy of the system when the particle is completely in the water phase.
We define $\Delta E_{I}$ as

\begin{equation}
    E_I = A_w \sigma_{pw} + A_o\sigma_{po} + A_{I_1}\sigma_{ow} + V_w P_w + V_o P_o
\end{equation}

Where subscripts $w$, $o$ and $I$ represents the water phase, oil phase and interface respectively. $A_w$ and $A_o$ are the surface areas of the particle in the water and oil phases 
respectively, $V_w$ and $V_o$ are the volumes of water and oil phases respectively, $P_w$ and $P_o$ are the pressures of water and oil phases respectively. $A_{I_1}$ is the area of the 
interface with the particle. A similar equation can be written for the energy of the water phase, $E_w$

\begin{equation}
    E_w = (A_w + A_o)\sigma_{pw} + A_{I_2}\sigma_{ow} + (V_w + V_o)P_w
\end{equation}

$A_{I_2}$ is the area of the interface without the particle. The change in the interface energy upon adsorption of the particle can be computed to obtain

\begin{equation}
    \Delta E_{IW} = A_o(\sigma_{po} - \sigma_{pw}) - A_p\sigma_{ow} + V_o(P_o - P_w)
\end{equation}

A similar construction can be done for the oil phase to obtain

\begin{equation}
    % \Delta E_{IO} = -\sigma_{ow}(A_w\cos{\theta_c} + A_p - \frac{2 V_o}{R_d})
    \Delta E_{IO} = A_w(\sigma_{pw} - \sigma_{po}) + A_p\sigma_{ow} - V_o(P_o - P_w)
\end{equation}

We now use the Young-Dupre equation relating the surface tensions of each component to one another, $\cos{\theta_c} = \frac{\sigma_{po} - \sigma_{pw}}{\sigma_{ow}}$ and the
Young-Laplace equation for a sphere, relating the pressure difference inside and outside a droplet $\Delta P$ to the radius of the droplet $R_d$, $\Delta P = \frac{2\sigma}{R_d}$
Substutions for $\sigma_{pw} - \sigma_{po}$ and $P_o - P_w$ are made, resulting in

\begin{equation}
    \begin{split}
        \Delta E_{IW} &= \sigma_{ow}(A_o\cos{\theta_c} - A_p - \frac{2 V_o}{R_d}) \\
        \Delta E_{IO} &= -\sigma_{ow}(A_w\cos{\theta_c} + A_p - \frac{2 V_o}{R_d})
    \end{split}
\end{equation}

From this the Pieranski model, the adsorption energy can be seen to depend on the contact angle, surface tension and contact area of the particles. This indicates that larger, neutrally wetting
particles adsorbing onto interfaces with high surface tension have a high driving force for adsorption. \cite{ngai_particle-stabilized_2015,reeves_particle-size_2015} This adsorption process 
can be orders of magnitude larger than the thermal energy of the system, calculated as $k_b T$ where $T$ is the temperature of the system and $k_b$ is the Boltzmann constant. This results in 
irreversible adsorption of particles at interfaces. \cite{binks_pickering_2001}

In addition to the adsorption of particles on the interface, the contact angle of the particles determine the dispersed phase of an emulsion and its microstructure. \cite{velankar_non-equilibrium_2015}
Hydrophillic particles tend to prefer to form oil in water emulsions while hydrophobic particles prefer to form water in oil emulsions. \cite{ngai_particle-stabilized_2015}
This occurs as the particles preferentially wet one phase over another, causing the interface to bend as the emulsion forms. Neutrally wetting particles are preferred in the formation of 
bijels as no preferential curvature is imparted onto the bijels, fulfilling one of the requirements of a minimal surface in that the mean curvature must be zero. \cite{jinnai_interfacial_2001}

@@

In contrast, explicit particle models represent colloids as finite-size rigid bodies with well-defined shape, volume, and surface boundary conditions
\cite{ladd_numerical_1994, jansen_bijels_2011,gunther_timescales_2014}. These particles interact 
with the surrounding fluid through no-slip or slip boundary conditions, and their motion responds dynamically to hydrodynamic, capillary, and interparticle forces. By resolving 
particle geometry, these models allow for accurate representation of rotational motion, surface wetting and interfacial adsorption
\cite{jansen_bijels_2011, gunther_lattice_2013}.
This is important in bijel systems, where particle-induced jamming depend on particle shape, size, and surface wetting.

An intermediate approach between point particles and explicitly resolved rigid bodies is the immersed boundary method (IBM), which represents extended particles as collections of 
point markers embedded in the fluid \cite{peskin_immersed_2002, luo_immersed-boundary_2008, spandan_parallel_2017}.
These markers interact with the fluid through momentum exchange that impose motion onto the surrounding fluid, effectively reconstructing 
the dynamics of a deformable or rigid object without requiring an explicit boundary mesh \cite{peskin_immersed_2002, luo_immersed-boundary_2008, spandan_parallel_2017}.
IBM has been widely used in biological and colloidal systems, to model the dynamics of deformable bodies \cite{peskin_immersed_2002, luo_immersed-boundary_2008, spandan_parallel_2017}.

For systems in which interfacial structure, jamming, or near-field hydrodynamics of rigid bodies play a central role as in bijel formation, explicit particle modeling 
is generally preferred as we are able to resolve the interfacial structure and near field hydrodynamics without the complications of IBM. This technique has been used in simulations of
bijels based on the Lattice Boltzmann Method and phase field techniques demonstrating its effectiveness in modeling
phenomena relevant to bijels \cite{jansen_bijels_2011,gunther_timescales_2014,carmack_tuning_2018}.

@@

Magnetically responsive particles can generally be classified as paramagnetic or ferromagnetic, depending on the nature and strength of their magnetization in response to external 
magnetic fields. Paramagnetic particles do not possess a permanent magnetic moment but acquire a reversible magnetization that aligns with the applied field and 
vanishes once the field is removed \cite{sinn_magnetically_2011}. Their response is typically linear with field strength and does not involve magnetic hysteresis. 
In contrast, ferromagnetic particles exhibit strong, permanent magnetization due to the alignment of magnetic domains even in the absence of an external field. These particles can 
exhibit hysteresis and significantly stronger dipole-dipole interactions \cite{sinn_magnetically_2011}.
As a result, ferromagnetic particles are more likely to form irreversible aggregates or chains, while paramagnetic particles 
respond more gently and reversibly, making them preferable in systems where tunability and field-off stability are important. The choice between these types depends on the desired 
degree of responsiveness, reversibility, and control in applications such as field-directed assembly or responsive emulsions.

% Magnetic particles can be modeled using the magnetic dipole approximation, which treats each particle as possessing a single point dipole moment \(\vec{m}\) located at its center
% \cite{davies_assembling_2014}.

% The dipole-dipole interaction between two particles separated by a vector \(\vec{r}\) is described by the potential:

% \begin{equation}
%     U_{\text{d}} = \frac{\mu_0}{4\pi |\vec{r}|^3} \left[ \vec{m}_1 \cdot \vec{m}_2 - 3(\vec{m}_1 \cdot \vec{r})(\vec{m}_2 \cdot \vec{r}) \right]
% \end{equation}
Magnetic particles can be modeled using the magnetic dipole approximation, which treats each particle as possessing a single point dipole moment \(\vec{m}\) located at its center
\cite{davies_assembling_2014}. 
This model captures the long-range, anisotropic nature of magnetic interactions and is commonly used due to its simplicity and computational efficiency. It is especially effective when 
particles are small, well-separated, and uniformly magnetized, such that higher-order multipole contributions can be neglected. However, the dipole model becomes less accurate when particles 
aggregate closely or experience strong magnetic fields that cause internal magnetization to deviate from a uniform state.

To address these limitations, the demagnetization tensor method offers a more refined model by considering the particle as a uniformly magnetized ellipsoid and calculating the internal magnetic 
field using an analytical or numerically tabulated demagnetization tensor \cite{takahashi_ellipsoids_2017}. 
This method accounts for the effect of particle shape on the distribution of internal magnetization and enables the modeling of induced magnetic 
dipoles that respond to local fields. This approach is particularly useful for modeling spheroidal or ellipsoidal particles and is widely 
used in magnetorheological systems, chain-forming ferrofluids, and dense suspensions where particle-particle interactions are shape-dependent.

% The total magnetic field of an ellipsoids can be written,

% \begin{equation}
%     \mathbf{H}(r) = \mathbf{H}_0 + \mathbf{N}(r)\mathbf{K}\mathbf{H}^{\dagger}
% \end{equation}

% Where $\mathbf{H}(r)$ is the total magnetic field of the ellipsoids, $\mathbf{H}_0$ is the inducing magnetic field in each direction,
% $\mathbf{K}$ is the susceptibility tensor and $\mathbf{H}^{\dagger}$ is the resultant uniform magnetic field at any point within
% the ellipsoids. \cite{takahashi_ellipsoids_2017}  This approach is particularly useful for modeling spheroidal or ellipsoidal particles and is widely 
% used in magnetorheological systems, chain-forming ferrofluids, and dense suspensions where particle-particle interactions are shape-dependent.

Since bijel formation and stabilization primarily depend on interfacial jamming and long-range interactions between particles at fluid-fluid interfaces, the magnetic dipole approximation 
is often sufficient to model key behaviors such as chaining, orientation under external fields, and capillary-mediated assembly. However, if incoroprating the effect of particle shape anisotropy, 
proximity effects, or induced magnetization in response to local fields are important, incorporating demagnetization tensors may offer improved realism. Investigations of particle reorientation at
interfaces has been conducted using the magnetic dipole model, showing that a rich variety of interfacial phenomena can be captured by this model such as the balance between capillary interactions
and magnetic field induced torque and capillary self assembly of microparticles at interfaces. \cite{xie_direct_2017,davies_interface_2014,davies_assembling_2014}
